---
title: "Pipeline for getting SFS data for sh deciles"
author: "Gustavo V. Barroso"
date: "`r Sys.Date()`"
output:
  pdf_document:
  toc: true
number_sections: true
toc_depth: 1
---

```{r setup, include=F, message=F, warning=F}
knitr::opts_chunk$set(echo=TRUE)

library(tidyverse)
library(data.table)
library(cowplot)
library(biomaRt)
library(reticulate)
library(plyr)
library(GenomicRanges)
library(vcfR)
library(foreach)
library(doParallel)

use_python("/usr/bin/python3")

knitr::opts_knit$set(root.dir="~/Data/bgs_lmr/human_data/data/sfs_genes")
```

```{bash, mk-dirs, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

mkdir masks
mkdir masks/bed
mkdir roulette_lof_deciles
mkdir roulette_lof_deciles/VEP_filtered
mkdir roulette_lof_deciles/roulette_annotated
mkdir roulette_lof_deciles/bed
mkdir sfs_output
```

## Getting gene information

Reads table with GeneBayes predictions and combines it with exonic coordinates

```{r, info_genes, include=T, eval=T, results=F, message=F, warning=T}

# https://www.nature.com/articles/s41588-024-01820-9
sh_table <- fread("other/ZengEtAl2024NatGen.csv")
names(sh_table)[1] <- "ensembl_gene_id"
sh_table <- filter(sh_table, chrom!="chrX", chrom!="chrY")

archives <- listEnsemblArchives()

ensembl_human <- useMart(biomart="ENSEMBL_MART_ENSEMBL", 
                         dataset="hsapiens_gene_ensembl",
                         host="https://oct2024.archive.ensembl.org")

ccds_genes <- getBM(mart=ensembl_human, 
                    filters=c("chromosome_name", "with_ccds"), 
                    values=list(c(1:22), T),
                    attributes=c("ensembl_gene_id", 
                                 "ensembl_exon_id",
                                 "ensembl_transcript_id",
                                 "chromosome_name",
                                 "transcript_start",
                                 "transcript_end",
                                 "exon_chrom_start", 
                                 "exon_chrom_end",
                                 "5_utr_start",
                                 "5_utr_end",
                                 "3_utr_start",
                                 "3_utr_end",
                                 "cds_start",
                                 "cds_end",
                                 "cds_length",
                                 "strand")) 

pred_genes <- getBM(mart=ensembl_human, 
                    filters=c("chromosome_name"), 
                    values=list(1:22),
                    attributes=c("ensembl_gene_id", 
                                 "ensembl_exon_id",
                                 "ensembl_transcript_id",
                                 "chromosome_name",
                                 "transcript_start",
                                 "transcript_end",
                                 "exon_chrom_start", 
                                 "exon_chrom_end",
                                 "5_utr_start",
                                 "5_utr_end",
                                 "3_utr_start",
                                 "3_utr_end",
                                 "cds_start",
                                 "cds_end",
                                 "cds_length",
                                 "strand")) 

fwrite(ccds_genes, "ccds_genes_hg38.csv")
fwrite(pred_genes, "pred_genes_hg38.csv")

ccds_genes <- fread("ccds_genes_hg38.csv")
pred_genes <- fread("pred_genes_hg38.csv")

# decides which set of genes to use
used_genes <- ccds_genes # ccds (VEP assigns too many MODIFIER sites otherwise)

# the goal is to have a BED file with positions that are CDS in SOME transcript
exon_singletons <- used_genes[!duplicated(used_genes$ensembl_exon_id),]

length(unique(exon_singletons$ensembl_gene_id))
length(unique(used_genes$ensembl_gene_id))

# focus on exons on the tail(s) of each gene
tails <- exon_singletons[!is.na(exon_singletons$`5_utr_start`) |
                         !is.na(exon_singletons$`3_utr_start`),]
tails_pos_strand <- filter(tails, strand==1) # positive strand 5'->3'
tails_neg_strand <- filter(tails, strand==-1) # negative strand 3'->5'

# clipping exon starts
tails_pos_strand[!is.na(tails_pos_strand$`5_utr_end`),]$exon_chrom_start <- 
  tails_pos_strand[!is.na(tails_pos_strand$`5_utr_end`),]$`5_utr_end`
tails_pos_strand[!is.na(tails_pos_strand$`3_utr_start`),]$exon_chrom_end <- 
  tails_pos_strand[!is.na(tails_pos_strand$`3_utr_start`),]$`3_utr_start`

tails_neg_strand[!is.na(tails_neg_strand$`3_utr_end`),]$exon_chrom_start <- 
  tails_neg_strand[!is.na(tails_neg_strand$`3_utr_end`),]$`3_utr_end`
tails_neg_strand[!is.na(tails_neg_strand$`5_utr_start`),]$exon_chrom_end <- 
  tails_neg_strand[!is.na(tails_neg_strand$`5_utr_start`),]$`5_utr_start`

# putting it back together
new_tails <- rbind.data.frame(tails_pos_strand, tails_neg_strand)
length(unique(new_tails$ensembl_exon_id)) == nrow(new_tails)

inner_exons <- filter(used_genes, !(ensembl_exon_id %in% new_tails$ensembl_exon_id))
unique_inner_exons <- filter(inner_exons, !duplicated(ensembl_exon_id))
clipped_genes <- rbind.data.frame(new_tails, unique_inner_exons)
clipped_genes <- filter(clipped_genes, exon_chrom_start!=exon_chrom_end)

nrow(clipped_genes) == length(unique(clipped_genes$ensembl_exon_id))
length(unique(clipped_genes$ensembl_gene_id))

clipped_genes <- dplyr::arrange(.data=clipped_genes, 
                                chromosome_name,
                                ensembl_gene_id, 
                                ensembl_transcript_id,
                                exon_chrom_start,
                                exon_chrom_end)

genes_sh <- merge(clipped_genes, sh_table, by="ensembl_gene_id")
genes_sh <- dplyr::arrange(.data=genes_sh, chromosome_name, exon_chrom_start)
sum(genes_sh$exon_chrom_end - genes_sh$exon_chrom_start)
length(unique(genes_sh$ensembl_gene_id))

# check if we are losing genes when merging with the Zeng et al table
length(unique(ccds_genes$ensembl_gene_id)) # ccds genes from ensembl
length(unique(clipped_genes$ensembl_gene_id)) # same, after trimming exon ends
length(unique(sh_table$ensembl_gene_id)) # zeng et al table for autosomes
length(unique(genes_sh$ensembl_gene_id)) # merged table

clipped_leftover <- dplyr::filter(clipped_genes, !ensembl_gene_id %in% genes_sh$ensembl_gene_id)
length(unique(clipped_leftover$ensembl_gene_id))

ccds_clipped_leftover <- filter(clipped_leftover, ensembl_gene_id %in% ccds_genes$ensembl_gene_id)
# 757 ccds genes not in the Zeng table
length(unique(ccds_clipped_leftover$ensembl_gene_id)) 

# including genes not in Zeng table as 11th decile, but only those with CCDS
ccds_clipped_leftover$hgnc <- NA
ccds_clipped_leftover$chrom <- NA
ccds_clipped_leftover$obs_lof <- NA
ccds_clipped_leftover$exp_lof <- NA
ccds_clipped_leftover$prior_mean <- NA
ccds_clipped_leftover$post_mean <- NA
ccds_clipped_leftover$post_lower_95 <- NA
ccds_clipped_leftover$post_upper_95 <- NA
ccds_clipped_leftover$decile_constraint <- 11 # 11th decile 

genes_sh_ext <- rbind.data.frame(genes_sh, ccds_clipped_leftover)
genes_sh_ext <- dplyr::arrange(.data=genes_sh_ext, 
                               chromosome_name, 
                               exon_chrom_start)
length(unique(genes_sh$ensembl_gene_id)) 
length(unique(genes_sh_ext$ensembl_gene_id)) # 18322 = 17565 + 757
genes_sh_ext <- filter(genes_sh_ext, exon_chrom_start!=exon_chrom_end)

sum(genes_sh$exon_chrom_end - genes_sh$exon_chrom_start)
sum(genes_sh_ext$exon_chrom_end - genes_sh_ext$exon_chrom_start)
sum(ccds_clipped_leftover$exon_chrom_end - ccds_clipped_leftover$exon_chrom_start)

# 724 genes in the Zeng et al table not in the ccds genes from Ensembl
zeng_leftover <- filter(sh_table, !ensembl_gene_id %in% genes_sh_ext$ensembl_gene_id)
nrow(zeng_leftover) # we will discard these, as there are no start, end coords.

fwrite(genes_sh_ext, "genes_sh_deciles_hg38.csv", sep=",")
genes_sh_ext <- fread("genes_sh_deciles_hg38.csv")

# sum of gene lengths per decile
cds_decile <- cbind.data.frame(rep(0, 11), rep(0, 11))
names(cds_decile) <- c("sum_cds", "num_genes")
cds_decile$decile <- 1:11

for(c in unique(genes_sh_ext$chromosome_name)) {
  for(d in unique(genes_sh_ext$decile_constraint)) {
    tbl <- dplyr::filter(genes_sh_ext, chromosome_name==c, decile_constraint==d)
    df <- dplyr::select(tbl, c("chromosome_name",
                               "exon_chrom_start",
                               "exon_chrom_end"))
    
    gr <- makeGRangesFromDataFrame(df, keep.extra.columns=T) # GR is 1-based
    df <- as.data.table(reduce(gr)) # remove overlaps
    df <- dplyr::select(df, c(seqnames, start, end))
    names(df) <- c("#chrom", "start", "end")
    
    # increments sum for each chr
    cds_decile$sum_cds[d] <- cds_decile$sum_cds[d] + sum(df$end - df$start) 
    cds_decile$num_genes[d] <- cds_decile$num_genes[d] + 
                               length(unique(tbl$ensembl_gene_id))
      
    # to 0-based, half-open intervals
    df$start <- df$start - 1 
    
    fwrite(df, paste("~/Data/bgs_lmr/human_data/data/sfs_genes/masks/bed/exons_",
                     "lof_d", d, "_chr", c, ".bed.gz", sep=""), sep="\t")
  }
}

plot(x=cds_decile$decile, y=cds_decile$num_genes,
     ylab="num genes", xlab="decile")
plot(x=cds_decile$decile, y=cds_decile$sum_cds / cds_decile$num_genes,
     ylab="avg gene lengths", xlab="decile")
```

## NYGC VCF files

NYGC high-coverage resequencing of 1kGP samples. 
Reference: Byrska-Bishop et al 2022, Cell

Downloaded on February 27 2025 and stored in:
/media/gvbarroso/extradrive1/bgs_lmr/human_data/data/1kGP

```{bash, vcf-download, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

for c in {1..22};
  wget https://ftp.1000genomes.ebi.ac.uk/vol1/ftp/data_collections/1000G_2504_high_coverage/working/20220422_3202_phased_SNV_INDEL_SV/1kGP_high_coverage_Illumina.chr$c.filtered.SNV_INDEL_SV_phased_panel.vcf.gz
done
```

We are keeping bi-allelic SNPs for YRI and CEU populations

```{bash, vcf-filter, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

for c in {1..22}; 
  bcftools view -v snps \
  1kGP_high_coverage_Illumina.chr$c.filtered.SNV_INDEL_SV_phased_panel.vcf.gz \
  -S ../yri_samples.txt | bgzip > 1kGP_YRI_filtered_SNPS_chr$c.vcf.gz
  bcftools view -v snps \
  1kGP_high_coverage_Illumina.chr$c.filtered.SNV_INDEL_SV_phased_panel.vcf.gz \
  -S ../ceu_samples.txt | bgzip > 1kGP_CEU_filtered_SNPS_chr$c.vcf.gz
done
```

## Roulette mutation maps (also VCF format)

Downloading and processing

```{bash, roulette-dl, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

for c in {1..22};
  wget http://genetics.bwh.harvard.edu/downloads/Vova/Roulette/${c}_rate_v5.2_TFBS_correction_all.vcf.bgz
  zcat ${c}_rate_v5.2_TFBS_correction_all.vcf.gz | \
  bgzip > chr${c}_rate_v5.2_TFBS_correction_all.vcf.gz
  tabix chr${c}_rate_v5.2_TFBS_correction_all.vcf.gz
  
  bcftools view -h chr${c}_rate_v5.2_TFBS_correction_all.vcf.gz | \
  sed '31i ##FILTER=<ID=high,Description="High quality variant; FILTER created by GVB on Apr 23 2025 to prevent errors when annotating the Roulette maps with VEP">' > hdr.txt
  bcftools reheader -h hdr.txt chr${c}_rate_v5.2_TFBS_correction_all.vcf.gz > \
  chr${c}_rate_v5.2_TFBS_correction_all_newheader.vcf.gz
  tabix chr${c}_rate_v5.2_TFBS_correction_all_newheader.vcf.gz
done
```

Intersecting Roulette map and exons in sh deciles (LOF/sh)

```{bash, roulette-mask, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

for c in {1..22}; do
  for d in {1..11}; do
    bcftools view /media/gvbarroso/Elements/Data/Roulette_hg38/vcf_files/chr${c}_rate_v5.2_TFBS_correction_all_newheader.vcf.gz \
    -R masks/bed/${c}_exons_lof_d${d}.tsv.gz | \
    bgzip > roulette_lof_deciles/roulette_chr${c}_lof_d${d}.vcf.gz;
  done
done

for f in roulette_lof_deciles/roulette_chr*; do tabix "$f"; done
```

Annotating Roulette maps

```{bash, roulette-mask, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

for c in {1..22}; do
  for d in {1..11}; do
    ./vep --cache -i roulette_lof_deciles/roulette_chr${c}_lof_d${d}.vcf.gz \
    -o roulette_lof_deciles/vep_roulette_chr${c}_lof_d${d}.tsv
    gzip roulette_lof_deciles/vep_roulette_chr${c}_lof_d${d}.tsv
  done
done

mkdir VEP_output
mv roulette_lof_deciles/vep_roulette_chr* VEP_output
```

Now we filter the VEP output by selecting the strongest consequence from each
possible mutation.

```{r, filter-vep-out, include=T, eval=T, results=F, warning=F}

# commented-out code perform decile loop in parallel, but uses a lot of memory
pb <- txtProgressBar(min=0, max=22, style=3)
for(c in 1:22) {
  setTxtProgressBar(pb, c)
  #cluster <- makeCluster(11)
  #registerDoParallel(cluster)
  #foreach(d = 1:11) %dopar% {
  for (d in 1:11) {
    fin <- paste("VEP_output/vep_roulette_chr", c, 
                 "_lof_d", d, ".tsv.gz", sep="")
    vep <- fread(paste("roulette_lof_deciles/", fin, sep=""), header=T)

    vep$POS <- as.integer(str_split_i(vep$`#Uploaded_variation`, 
                                      "_", 2))
    vep$REF <- str_split_i(str_split_i(vep$`#Uploaded_variation`, 
                                       "_", 3), "/", 1)
    vep$VEP <- str_split_i(str_split_i(vep$Extra, ";", 1), "=", 2)
    vep$CHROM <- c
    vep <- dplyr::select(vep, c(CHROM, POS, REF, Allele, Gene, Feature, VEP))
    vep$Score <- as.integer(mapvalues(vep$VEP, c("MODIFIER", 
                                                 "LOW", 
                                                 "MODERATE", 
                                                 "HIGH"), 1:4))
    
    vep_filtered <- vep %>%
                    group_by(POS, Allele) %>% 
                    dplyr::filter(., Score==max(Score)) %>%
                    distinct(POS, Allele, .keep_all=TRUE)
    
    table(vep$VEP) # prior to filtering and uniquing
    table(vep_filtered$VEP) # after filtering and uniquing
    
    names(vep_filtered)[which(names(vep_filtered)=="Allele")] <- "ALT"
    vep_filtered <- dplyr::select(vep_filtered, c(CHROM, POS, REF, ALT, VEP))
    
    vep_high <- filter(vep_filtered, VEP=="HIGH") # LOF
    vep_mod <- filter(vep_filtered, VEP=="MODERATE") # nonsynonymous
    vep_low <- filter(vep_filtered, VEP=="LOW") # synonymous

    # writing uncompressed files as they will be bgzip-ed
    fout <- paste("vep_LOW_chr", c, "_d", d, ".tsv", sep="")
    fwrite(vep_low, paste("roulette_lof_deciles/VEP_filtered/", fout, sep=""),
           sep="\t", col.names=F)
    
    fout <- paste("vep_MODERATE_chr", c, "_d", d, ".tsv", sep="")
    fwrite(vep_mod, paste("roulette_lof_deciles/VEP_filtered/", fout, sep=""),
           sep="\t", col.names=F)
    
    fout <- paste("vep_HIGH_chr", c, "_d", d, ".tsv", sep="")
    fwrite(vep_high, paste("roulette_lof_deciles/VEP_filtered/", fout, sep=""),
           sep="\t", col.names=F)
  }
  # recycle
  #stopCluster(cl = cluster)
}
close(pb)
```

Adding INFO field to Roulette VCFs based on VEP impact

```{bash, annotate-roulette, engine.opts='-i', include=T, eval=T, results=F, warning=F}

source ~/.bashrc

cd roulette_lof_deciles/VEP_filtered
echo "##INFO=<ID=VEP,Number=1,Type=String,Description=VEP-predicted impact" \
     > header.txt
for f in vep_*; do bgzip $f; done
for f in *.tsv.gz; do tabix -s1 -b2 -e2 $f; done
cd ..

for c in {1..22}; do 
  for d in {1..11}; do 
    bcftools annotate \
    -a VEP_filtered/vep_HIGH_chr${c}_d${d}.tsv.gz \
    roulette_chr${c}_lof_d${d}.vcf.gz \
    -c CHROM,POS,REF,ALT,VEP \
    -h VEP_filtered/header.txt | bgzip -c > \
    roulette_annotated/roulette_chr${c}_lof_d${d}_vep_HIGH.vcf.gz
    
    bcftools annotate \
    -a VEP_filtered/vep_MODERATE_chr${c}_d${d}.tsv.gz \
    roulette_chr${c}_lof_d${d}.vcf.gz \
    -c CHROM,POS,REF,ALT,VEP \
    -h VEP_filtered/header.txt | bgzip -c > \
    roulette_annotated/roulette_chr${c}_lof_d${d}_vep_MODERATE.vcf.gz
    
    bcftools annotate \
    -a VEP_filtered/vep_LOW_chr${c}_d${d}.tsv.gz \
    roulette_chr${c}_lof_d${d}.vcf.gz \
    -c CHROM,POS,REF,ALT,VEP \
    -h VEP_filtered/header.txt | bgzip -c > \
    roulette_annotated/roulette_chr${c}_lof_d${d}_vep_LOW.vcf.gz
  done
done

for f in roulette_annotated/roulette_chr*; do tabix "$f"; done
for f in roulette_annotated/roulette_chr*; do bcftools view "$f"; done # ?

cd roulette_annotated

# stores headers as separate files (not used in the current version of pipeline)
for c in {1..22}; do 
  for d in {1..11}; do 
    bcftools view -h roulette_chr${c}_lof_d${d}_vep_LOW.vcf.gz > \
    roulette_chr${c}_lof_d${d}_vep_LOW_header.txt
    
    bcftools view -h roulette_chr${c}_lof_d${d}_vep_MODERATE.vcf.gz > \
    roulette_chr${c}_lof_d${d}_vep_MODERATE_header.txt
    
    bcftools view -h roulette_chr${c}_lof_d${d}_vep_HIGH.vcf.gz > \
    roulette_chr${c}_lof_d${d}_vep_HIGH_header.txt
  done
done

cd ..
```

Filtering roulette VCFs based on VEP impact to get U and L per decile per impact

```{r, filter-roulette, include=T, eval=T, results=F, warning=F}

# https://github.com/vseplyarskiy/Roulette/tree/main/adding_mutation_rate
roulette_scale <- 1.015e-7 / 2 # factor of 2 because it's per diploid genome

vep_classes <- c("LOW", "MODERATE", "HIGH")

L_d <- cbind.data.frame(rep(0, 11), rep(0, 11), rep(0, 11))
names(L_d) <- vep_classes

U_d <- cbind.data.frame(rep(0, 11), rep(0, 11), rep(0, 11))
names(U_d) <- vep_classes

# flip the order of the nested loop to get avg mutation rate per decile
pb <- txtProgressBar(min=0, max=11, style=3)
for(d in 1:11) { 

  setTxtProgressBar(pb, d)
  
  mus_d <- c(0, 0, 0) 
  names(mus_d) <- vep_classes
  
  nsites_d <- c(0, 0, 0) 
  names(nsites_d) <- vep_classes
  
  for(i in 1:length(vep_classes)) { 
    ip <- vep_classes[i]
    pattern <- paste("VEP", ip, sep="=")

    idx <- which(names(mus_d) == ip)
    
    for(c in 1:22) { 
      vcf <- fread(paste("roulette_lof_deciles/roulette_annotated/",
                          paste("roulette_chr", c, 
                                 "_lof_d", d, 
                                 "_vep_", ip, 
                                 ".vcf.gz", sep=""), sep="")) 
      
      vcf <- dplyr::filter(vcf, grepl(pattern, INFO))
  
      # gets roulette mutation rates 
      mr <- str_split_i(vcf$INFO, ";", 2)
      mr <- as.numeric(str_split_i(mr, "=", 2)) * roulette_scale
      mus_d[idx] <- mus_d[idx] + sum(mr) # mu * L
      
      nsites_d[idx] <- nsites_d[idx] + length(unique(vcf$POS)) # L
    }
  }
  
  L_d[d,] <- nsites_d
  U_d[d,] <- mus_d
}
close(pb)

# mu * L and L, summed across chromosomes
fwrite(U_d, "~/Data/bgs_lmr/human_data/data/sfs_genes/sfs_output/U_d.csv") 
fwrite(L_d, "~/Data/bgs_lmr/human_data/data/sfs_genes/sfs_output/L_d.csv") 
```

We now filter NYG VCFs based on predicted impact of REF->ALT mutation

```{r, filter-NYG-VCFs, include=T, eval=T, results=F, warning=F}

for(c in 1:22) {
  cat(paste("chromosome", c, "\n"))
  
  data_1kgp <- fread(paste("/media/gvbarroso/extradrive1/bgs_lmr/human_data/data/1kGP/",
                     paste("1kGP_CEU_filtered_SNPS_chr", c, ".vcf.gz", sep=""), 
                     sep=""), skip=112) # %>% dplyr::select(c(1,2,4,5))
  
  for(d in 1:11) {
    cat(paste("decile", d, "\n"))
    
    # selecting only columns we need to intersect
    rlt_vep_L <- fread(paste("roulette_lof_deciles/roulette_annotated/",
                             paste("roulette_chr", c, 
                                   "_lof_d", d, 
                                   "_vep_LOW", 
                                   ".vcf.gz", sep=""), sep="")) %>%
                 dplyr::filter(., grepl("VEP=LOW", INFO)) %>%
                 dplyr::select(., c(POS, REF, ALT)) 
    
    rlt_vep_M <- fread(paste("roulette_lof_deciles/roulette_annotated/",
                             paste("roulette_chr", c, 
                                   "_lof_d", d, 
                                   "_vep_MODERATE", 
                                   ".vcf.gz", sep=""), sep="")) %>%
                 dplyr::filter(., grepl("VEP=MODERATE", INFO)) %>%
                 dplyr::select(., c(POS, REF, ALT))
    
    rlt_vep_H <- fread(paste("roulette_lof_deciles/roulette_annotated/",
                             paste("roulette_chr", c, 
                                   "_lof_d", d, 
                                   "_vep_HIGH", 
                                   ".vcf.gz", sep=""), sep="")) %>%
                 dplyr::filter(., grepl("VEP=HIGH", INFO)) %>%
                 dplyr::select(., c(POS, REF, ALT)) 
          
    vcf_L <- inner_join(data_1kgp, rlt_vep_L, by=c("POS", "REF", "ALT"))
    vcf_M <- inner_join(data_1kgp, rlt_vep_M, by=c("POS", "REF", "ALT"))
    vcf_H <- inner_join(data_1kgp, rlt_vep_H, by=c("POS", "REF", "ALT"))
    
    # writing headless VCF files
    fwrite(vcf_L, paste("~/Data/bgs_lmr/human_data/data/sfs_genes/filtered_vcfs/",
                        "1kGP_CEU_chr", c, "_d", d, "_VEP_LOW.vcf.gz", sep=""),
           sep="\t")
    
    fwrite(vcf_M, paste("~/Data/bgs_lmr/human_data/data/sfs_genes/filtered_vcfs/",
                        "1kGP_CEU_chr", c, "_d", d, "_VEP_MODERATE.vcf.gz", sep=""),
           sep="\t")
    
    fwrite(vcf_H, paste("~/Data/bgs_lmr/human_data/data/sfs_genes/filtered_vcfs/",
                        "1kGP_CEU_chr", c, "_d", d, "_VEP_HIGH.vcf.gz", sep=""),
           sep="\t")
    
    low <- vcf_L$POS
    mid <- vcf_M$POS
    high <- vcf_H$POS
    
    # to have an idea of the intersection that results from positions being
    # listed twice in the VCF files (although some are invariant)
    cat(paste(length(intersect(low, mid))), "-")
    cat(paste(length(intersect(low, high))), "-")
    cat(paste(length(intersect(mid, high))), "-")
    cat("\n")
  }
}
```

This python chunk was run as a script from outside this Rmd file

```{python, compute-SFS, engine.opts='-i', include=T, eval=T, results=F, warning=F}

import moments

for c in range(1,23):
  for d in range(1,12):
    for v in ["LOW", "MODERATE", "HIGH"]:
      for pop in ["YRI", "CEU"]:
        fout=f"sfs_output/sfs_{pop}_chr{c}_d{d}_VEP_{v}.sfs"
        vcfin=f"/media/gvbarroso/extradrive1/bgs_lmr/human_data/data/1kGP/1kGP_{pop}_filtered_SNPS_chr{c}.vcf.gz"
        bedin=f"~/Data/bgs_lmr/human_data/data/sfs_genes/roulette_lof_deciles/bed/chr{c}_lof_d{d}_VEP_{v}.bed"
        sfs=moments.Spectrum.from_vcf(vcfin, bed_file=bedin, folded=True)
        sfs.to_file(fout)
```
